---
title: "Coyotes_SpatialData_Jan2022"
author: "Jonathan Farr"
date: "1/24/2022"
output: html_document
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
```


# Loading packages
```{r}
library(raster)
library(sp)
library(arcgisbinding)
arc.check_product()
library(tidyverse)
```


# Spatial Data

## Reading in Coyote Points
```{r}
# read in coyote data from ArcGIS
Coyotes_arc = arc.select(arc.open("Data/GIS_Layers/UrbanCoyotesArcGIS/UrbanCoyotes.gdb/JJFCoyotesMarch82021_Projected"))

Coyotes.sp <- arc.data2sp(Coyotes_arc) # converting to spatial points object


```


## Reading in Raster files
With each line of code we are: 
a) reading in a raster from a specific folder
b) dividing the values in that raster by the max # cells in a buffer zone to generate proportions
c) using the extract function to assign those raster values to points in Coyotes.sp
d) adding these points to the correct data frame as a new column 

```{r}
# LAND COVER 100M BUFFER
Coyotes = data.frame(x = c(1:7878))
  
  
Coyotes$Com100m = raster::extract(raster("Data/GIS_Layers/LandCover100m/Com100m.tif")/317, Coyotes.sp)
Coyotes$Res100m = raster::extract(raster("Data/GIS_Layers/LandCover100m/Res100m.tif")/317, Coyotes.sp)
Coyotes$Mod100m = raster::extract(raster("Data/GIS_Layers/LandCover100m/Mod100m.tif")/317, Coyotes.sp)
Coyotes$Mow100m = raster::extract(raster("Data/GIS_Layers/LandCover100m/Mow100m.tif")/317, Coyotes.sp)
Coyotes$Nat100m = raster::extract(raster("Data/GIS_Layers/LandCover100m/Nat100m.tif")/317, Coyotes.sp)
Coyotes$Water100m = raster::extract(raster("Data/GIS_Layers/LandCover100m/Water100m.tif")/317, Coyotes.sp)
Coyotes$Build100m = raster::extract(raster("Data/GIS_Layers/LandCover100m/Build100m.tif")/317, Coyotes.sp)
Coyotes$Roads100m = raster::extract(raster("Data/GIS_Layers/LandCover100m/Roads100m.tif")/317, Coyotes.sp)

# LAND COVER 200M BUFFER
Coyotes$Com200m = raster::extract(raster("Data/GIS_Layers/LandCover200m/Com200m.tif")/1257, Coyotes.sp)
Coyotes$Res200m = raster::extract(raster("Data/GIS_Layers/LandCover200m/Res200m.tif")/1257, Coyotes.sp)
Coyotes$Mod200m = raster::extract(raster("Data/GIS_Layers/LandCover200m/Mod200m.tif")/1257, Coyotes.sp)
Coyotes$Mow200m = raster::extract(raster("Data/GIS_Layers/LandCover200m/Mow200m.tif")/1257, Coyotes.sp)
Coyotes$Nat200m = raster::extract(raster("Data/GIS_Layers/LandCover200m/Nat200m.tif")/1257, Coyotes.sp)
Coyotes$Water200m = raster::extract(raster("Data/GIS_Layers/LandCover200m/Water200m.tif")/1257, Coyotes.sp)
Coyotes$Build200m = raster::extract(raster("Data/GIS_Layers/LandCover200m/Build200m.tif")/1257, Coyotes.sp)
Coyotes$Roads200m = raster::extract(raster("Data/GIS_Layers/LandCover200m/Roads200m.tif")/1257, Coyotes.sp)

# LAND COVER 400M BUFFER
Coyotes$Com400m = raster::extract(raster("Data/GIS_Layers/LandCover400m/Com400m.tif")/5025, Coyotes.sp)
Coyotes$Res400m = raster::extract(raster("Data/GIS_Layers/LandCover400m/Res400m.tif")/5025, Coyotes.sp)
Coyotes$Mod400m = raster::extract(raster("Data/GIS_Layers/LandCover400m/Mod400m.tif")/5025, Coyotes.sp)
Coyotes$Mow400m = raster::extract(raster("Data/GIS_Layers/LandCover400m/Mow400m.tif")/5025, Coyotes.sp)
Coyotes$Nat400m = raster::extract(raster("Data/GIS_Layers/LandCover400m/Nat400m.tif")/5025, Coyotes.sp)
Coyotes$Water400m = raster::extract(raster("Data/GIS_Layers/LandCover400m/Water400m.tif")/5025, Coyotes.sp)
Coyotes$Build400m = raster::extract(raster("Data/GIS_Layers/LandCover400m/Build400m.tif")/5025, Coyotes.sp)
Coyotes$Roads400m = raster::extract(raster("Data/GIS_Layers/LandCover400m/Roads400m.tif")/5025, Coyotes.sp)

# LAND COVER 800M BUFFER
Coyotes$Com800m = raster::extract(raster("Data/GIS_Layers/LandCover800m/Com800m.tif")/20081, Coyotes.sp)
Coyotes$Res800m = raster::extract(raster("Data/GIS_Layers/LandCover800m/Res800m.tif")/20081, Coyotes.sp)
Coyotes$Mod800m = raster::extract(raster("Data/GIS_Layers/LandCover800m/Mod800m.tif")/20081, Coyotes.sp)
Coyotes$Mow800m = raster::extract(raster("Data/GIS_Layers/LandCover800m/Mow800m.tif")/20081, Coyotes.sp)
Coyotes$Nat800m = raster::extract(raster("Data/GIS_Layers/LandCover800m/Nat800m.tif")/20081, Coyotes.sp)
Coyotes$Water800m = raster::extract(raster("Data/GIS_Layers/LandCover800m/Water800m.tif")/20081, Coyotes.sp)
Coyotes$Build800m = raster::extract(raster("Data/GIS_Layers/LandCover800m/Build800m.tif")/20081, Coyotes.sp)
Coyotes$Roads800m = raster::extract(raster("Data/GIS_Layers/LandCover800m/Roads800m.tif")/20081, Coyotes.sp)

# LAND COVER 1600M BUFFER
Coyotes$Com1600m = raster::extract(raster("Data/GIS_Layers/LandCover1600m/Com1600m.tif")/80425, Coyotes.sp)
Coyotes$Res1600m = raster::extract(raster("Data/GIS_Layers/LandCover1600m/Res1600m.tif")/80425, Coyotes.sp)
Coyotes$Mod1600m = raster::extract(raster("Data/GIS_Layers/LandCover1600m/Mod1600m.tif")/80425, Coyotes.sp)
Coyotes$Mow1600m = raster::extract(raster("Data/GIS_Layers/LandCover1600m/Mow1600m.tif")/80425, Coyotes.sp)
Coyotes$Nat1600m = raster::extract(raster("Data/GIS_Layers/LandCover1600m/Nat1600m.tif")/80425, Coyotes.sp)
Coyotes$Water1600m = raster::extract(raster("Data/GIS_Layers/LandCover1600m/Water1600m.tif")/80425, Coyotes.sp)
Coyotes$Build1600m = raster::extract(raster("Data/GIS_Layers/LandCover1600m/Build1600m.tif")/80425, Coyotes.sp)
Coyotes$Roads1600m = raster::extract(raster("Data/GIS_Layers/LandCover1600m/Roads1600m.tif")/80425, Coyotes.sp)


Coyotes[is.na(Coyotes)] = 0 # raplacing all NAs with 0 

# Cleaning up the environment
Coyote_Data = cbind(Coyotes, as.data.frame(Coyotes.sp))


rm(Coyotes_arc, Coyotes.sp, Coyotes)
```

# Cleaning Coyote Data
```{r}
Coyotes = Coyote_Data %>% 
  mutate(Date = as.Date(Date, format = "%m/%d/%Y")) %>%
  mutate(covid = ifelse(Date > "2020-04-01", "Y", "N"), 
         CoyoteResponse = as.ordered(CoyoteResponse), 
         HumanPerception = as.ordered(HumanPerception))%>%
  filter(Date > "2010-12-31")
```


# Running some model real quick to make sure nothing changed in a big way 
```{r}
library(ordinal)

response = Coyotes %>%
  dplyr::filter(CoyoteResponse != "Unknown", CoyoteResponse != 0)

str(response)
response_full = clm(CoyoteResponse ~  Mod200m  + Nat200m +Mow200m + Com200m + Res200m + Season + as.factor(Year), data = response, na.action="na.fail")
summary(response_full)
dredge_response <- MuMIn::dredge(response_full) 
dredge_results_response <- MuMIn::model.avg(dredge_response, subset=delta < 2)

summary(clm(CoyoteResponse ~  Mod200m + Nat200m  +Mow200m + Com200m + Res200m, # + Season + as.factor(Year), 
            data = response, na.action="na.fail"))
# RES IS WEIRD - CHECK THIS OUT 
summary(dredge_results_response)
str(response)

```



```{r}
pup = subset(response, Season == "PupRearing")
breed= subset(response, Season == "Breeding")
dis = subset(response, Season == "Dispersal")

summary(clm(CoyoteResponse ~ RoadDistDecay + Mod200m  + Nat200m +Mow200m + Com200m  + as.factor(Year), data = pup, na.action="na.fail"))

summary(clm(CoyoteResponse ~ RoadDistDecay + Mod200m  + Nat200m +Mow200m + Com200m + as.factor(Year), data = breed, na.action="na.fail"))

summary(clm(CoyoteResponse ~RoadDistDecay + Mod200m  + Nat200m +Mow200m + Com200m+ as.factor(Year), data = dis, na.action="na.fail"))


```


















Potential alternate way to deal with it - just not sure how to cope with list objects going forward
```{r}
#first import all files in a single folder as a list 
rastlist <- list.files(path = "Data/GIS_Layers/LandCover100m", pattern='.tif$', 
all.files=TRUE, full.names=TRUE)

#import all raster files in folder using lapply
allrasters <- lapply(rastlist, raster)
allrastersProp = lapply(allrasters, function(x) x/317)

#to check the index numbers of all imported raster list elements
allrasters

#I want a loop that takes each raster, applies the "extract" function using the coyotes.sp points, and then puts that #data into a data frame that I can perhaps add as a column to the Coyotes.sp data


xy  = data.frame(raster::extract(allrastersProp[[1]], Coyotes.sp), 
                 raster::extract(allrastersProp[[2]], Coyotes.sp), 
                 raster::extract(allrastersProp[[3]], Coyotes.sp),
                 raster::extract(allrastersProp[[4]], Coyotes.sp),
                 raster::extract(allrastersProp[[5]], Coyotes.sp),
                 raster::extract(allrastersProp[[6]], Coyotes.sp),
                 raster::extract(allrastersProp[[7]], Coyotes.sp),
                 raster::extract(allrastersProp[[8]], Coyotes.sp))
            
colnames(xy) = c(allrastersProp[[1]]@data@names, allrastersProp[[2]]@data@names,
                              allrastersProp[[3]]@data@names, allrastersProp[[4]]@data@names,
                              allrastersProp[[5]]@data@names, allrastersProp[[6]]@data@names,
                              allrastersProp[[7]]@data@names, allrastersProp[[8]]@data@names)             




```

